---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# covid19viz

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/covid19viz)](https://cran.r-project.org/package=covid19viz)
[![DOI](https://zenodo.org/badge/246721190.svg)](https://zenodo.org/badge/latestdoi/246721190)
<!-- badges: end -->

The goal of `covid19viz` is to access and summarize WHO sitreps for covid-19 in simple graphics.

This package works using two data repositories:

- digitalized __WHO sitreps__ by [Fabienne Krauer](https://twitter.com/FabiKrauer) available in [fkrauer/COVID-19](https://github.com/fkrauer/COVID-19).

- Johns Hopkins University __(JHU CSSE)__ available in [CSSEGISandData/COVID-19](https://github.com/CSSEGISandData/COVID-19).

Additionally, it imports external data sources like:

- __Google__ Mobility Reports. [View](https://www.google.com/covid19/mobility/).

- __ACAPS__ data on #COVID19 Government Measures Dataset. [View](https://www.acaps.org/covid19-government-measures-dataset).

- __UNESCO__ COVID-19 Impact on Education. [View](https://en.unesco.org/covid19/educationresponse).

## Installation

<!--
You can install the released version of covid19viz from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("covid19viz")
```
-->

You can install the development version of `covid19viz` using:

``` r
if(!require("remotes")) install.packages("remotes")
remotes::install_github("avallecam/covid19viz")
```

## Quick Examples

### Import and visualize JHU collection

```{r}
library(covid19viz)
library(tidyverse)

#import all data at once
jhu_sitrep_peru <- jhu_sitrep_all_sources(country_region="Peru")

jhu_sitrep_peru

#transform to tidy format
jhu_sitrep_peru %>% 
  jhu_sitrep_all_sources_tidy() %>% 
  arrange(desc(confirmed_cumulative)) %>% 
  glimpse()
```

```{r,fig.width=8,fig.height=4,fig.align='center'}
jhu_sitrep_country_report(country_region = "Peru")
```

### Import more external data sources 

- Google Mobility Reports. [View](https://www.google.com/covid19/mobility/).
- ACAPS data on #COVID19 Government Measures Dataset. [View](https://www.acaps.org/covid19-government-measures-dataset).
- UNESCO COVID-19 Impact on Education. [View](https://en.unesco.org/covid19/educationresponse).

```{r,eval=FALSE}
library(covid19viz)
library(tidyverse)

# google mobility reports
# global
read_google_global() %>% glimpse()
#> Rows: 2,441,517
#> Columns: 14
#> $ country_region_code                                <chr> "AE", "AE", "AE", ...
#> $ country_region                                     <chr> "United Arab Emira...
#> $ sub_region_1                                       <chr> NA, NA, NA, NA, NA...
#> $ sub_region_2                                       <lgl> NA, NA, NA, NA, NA...
#> $ metro_area                                         <lgl> NA, NA, NA, NA, NA...
#> $ iso_3166_2_code                                    <chr> NA, NA, NA, NA, NA...
#> $ census_fips_code                                   <lgl> NA, NA, NA, NA, NA...
#> $ date                                               <date> 2020-02-15, 2020-...
#> $ retail_and_recreation_percent_change_from_baseline <dbl> 0, 1, -1, -2, -2, ...
#> $ grocery_and_pharmacy_percent_change_from_baseline  <dbl> 4, 4, 1, 1, 0, 1, ...
#> $ parks_percent_change_from_baseline                 <dbl> 5, 4, 5, 5, 4, 6, ...
#> $ transit_stations_percent_change_from_baseline      <dbl> 0, 1, 1, 0, -1, 1,...
#> $ workplaces_percent_change_from_baseline            <dbl> 2, 2, 2, 2, 2, 1, ...
#> $ residential_percent_change_from_baseline           <dbl> 1, 1, 1, 1, 1, 1, ...

# regional
# first: select country ISO
read_google_region_list()
#> # A tibble: 135 x 4
#>     country_iso Name                                 Length Date               
#>     <chr>       <chr>                                 <dbl> <dttm>             
#>   1 AE          2020_AE_Region_Mobility_Report.csv   121354 2020-09-13 15:27:00
#>   2 AF          2020_AF_Region_Mobility_Report.csv    20553 2020-09-13 15:27:00
#>   3 AG          2020_AG_Region_Mobility_Report.csv    57484 2020-09-13 15:27:00
#>   4 AO          2020_AO_Region_Mobility_Report.csv    58954 2020-09-13 15:27:00
#>   5 AR          2020_AR_Region_Mobility_Report.csv  6932241 2020-09-13 15:27:00
#>   6 AT          2020_AT_Region_Mobility_Report.csv  1402047 2020-09-13 15:27:00
#>   7 AU          2020_AU_Region_Mobility_Report.csv  4329094 2020-09-13 15:27:00
#>   8 AW          2020_AW_Region_Mobility_Report.csv     9967 2020-09-13 15:27:00

# second: read specific country data
peru <- read_google_region_country(country_iso = "PE")
peru %>% glimpse()
#> Rows: 28,194
#> Columns: 14
#> $ country_region_code                                <chr> "PE", "PE", "PE", ...
#> $ country_region                                     <chr> "Peru", "Peru", "P...
#> $ sub_region_1                                       <chr> NA, NA, NA, NA, NA...
#> $ sub_region_2                                       <chr> NA, NA, NA, NA, NA...
#> $ metro_area                                         <lgl> NA, NA, NA, NA, NA...
#> $ iso_3166_2_code                                    <chr> NA, NA, NA, NA, NA...
#> $ census_fips_code                                   <lgl> NA, NA, NA, NA, NA...
#> $ date                                               <date> 2020-02-15, 2020-...
#> $ retail_and_recreation_percent_change_from_baseline <dbl> 4, 1, 0, 0, 0, 0, ...
#> $ grocery_and_pharmacy_percent_change_from_baseline  <dbl> 1, 0, 1, 0, -1, -1...
#> $ parks_percent_change_from_baseline                 <dbl> 3, -2, 1, 0, 0, -1...
#> $ transit_stations_percent_change_from_baseline      <dbl> 3, 2, 2, 1, 0, 3, ...
#> $ workplaces_percent_change_from_baseline            <dbl> 0, 0, 1, 1, 1, 0, ...
#> $ residential_percent_change_from_baseline           <dbl> -1, 0, 0, 0, 1, 0,...

peru %>% count(sub_region_1)
#> # A tibble: 27 x 2
#>    sub_region_1      n
#>    <chr>         <int>
#>  1 Amazonas       1050
#>  2 Ancash         1871
#>  3 Apurimac        816
#>  4 Arequipa       1470
#>  5 Ayacucho       1050
#>  6 Cajamarca      1850
#>  7 Callao Region   210
#>  8 Cusco          1668
#>  9 Huancavelica    630
#> 10 Huanuco         630
#> # ... with 17 more rows

# acaps data
read_acaps_governments() %>% glimpse()
#> Rows: 193
#> Columns: 8
#> $ Country_name         <chr> "Afghanistan", "Albania", "Algeria", "Angola",...
#> $ ISO3                 <chr> "AFG", "ALB", "DZA", "AGO", "ATG", "ARG", "ARM...
#> $ REGION               <chr> "Asia", "Europe", "Africa", "Africa", "America...
#> $ Measure_taxonomy     <chr> "Schools closure ", "Border checks ", "Border ...
#> $ Measure_Category     <chr> "Social distancing", "Movement restrictions", ...
#> $ SOURCE_TYPE          <chr> "Government", "Media", "UN", "Other organisati...
#> $ Enforcement_taxonomy <chr> "Fines", "Arrest/Detention", "Deportation", "R...
#> $ Log_type             <chr> "Introduction / extension of measures", "Phase...

# unesco education data
read_unesco_education() %>% glimpse()
#> Rows: 41,714
#> Columns: 5
#> $ Date    <chr> "17/02/2020", "17/02/2020", "18/02/2020", "...
#> $ ISO     <chr> "CHN", "MNG", "CHN", "MNG", "CHN", "MNG", "...
#> $ Country <chr> "China", "Mongolia", "China", "Mongolia", "...
#> $ Status  <chr> "Partially open", "Closed due to COVID-19",...
#> $ Note    <lgl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,...
```

```{r,echo=FALSE,fig.height=4,fig.width=10,fig.retina=TRUE}
# library(covidPeru); library(readr)
library(covid19viz)
# library(cdcper)
library(tidyverse)
library(lubridate)
theme_set(theme_bw())


# analysis time limits --------------------------------------------------------------

min_analysis_date <- ymd(20200301)
max_analysis_date <- Sys.Date()

# covid19viz R package ----------------------------------------------------

# _import intervention data ------------------------------------------------

unesco <- read_unesco_education()
# acaps <- read_acaps_governments()

# no data in ACAPS
# acaps %>% filter(ISO3=="PER")

unesco_peru <- unesco %>% 
  filter(ISO=="PER") %>% 
  mutate(Date=dmy(Date)) %>% 
  group_by(Status) %>% 
  summarise(date_min=min(Date),
            date_max=max(Date)) %>% 
  rename(intervention_label=Status) %>% 
  mutate(intervention=case_when(
    intervention_label=="Closed due to COVID-19" ~ "closed",
    intervention_label=="Partially open" ~ "partial"
  ))
# unesco_peru

# _unite intervention data -------------------------------------------------

interventions <- tibble(
  date_min = ymd(20200628), 
  date_max = ymd(20200709),
  intervention_label = "Seroprevalence study",
  intervention = "seroprev"
) %>% 
  union_all(
    unesco_peru %>% 
      mutate(date_max=if_else(date_max==max(date_max),
                              max_analysis_date,
                              date_max))
  )
# interventions

# mobility data -----------------------------------------------------------

mobility <- read_google_region_country(country_iso = "PE")
# mobility %>% count(sub_region_1) %>% print(n=Inf)
# mobility %>% count(country_region,sub_region_1,sub_region_2,metro_area) %>% print(n=Inf)
mobility_lima <- mobility %>% 
  filter(date<max_analysis_date) %>% 
  filter(
    magrittr::is_in(sub_region_1,c("Metropolitan Municipality of Lima",
                                   "Callao Region"))) %>% 
  mutate(across(c(sub_region_1,sub_region_2,metro_area),
                str_replace_na,replacement = "")) %>% 
  mutate(subregion=str_c(sub_region_1,"\n",sub_region_2,"\n",metro_area)) %>% 
  mutate(subregion=str_trim(subregion)) %>% 
  mutate(subregion=if_else(sub_region_1=="Metropolitan Municipality of Lima",
                           "Metropolitan Municipality\nof Lima",subregion)) %>% 
  # count(country_region,subregion)
  filter(!(sub_region_1=="Metropolitan Municipality of Lima" &
             sub_region_2=="")) %>%
  # count(country_region,sub_region_1,sub_region_2,metro_area,subregion)
  # pivot_longer()
  pivot_longer(cols = -c(country_region_code:date,subregion),
               names_to = "field",
               names_pattern = "(.+)_percent_change_from_baseline",
               values_to = "percent_change_from_baseline")

# plot all --------------------------------------

mobility_lima %>% 
  mutate(field=str_replace_all(field,"_"," "),
         field=str_to_sentence(field)) %>% 
  ggplot() +
  geom_rect(data = interventions,
            aes(xmin = date_min, xmax = date_max, 
                ymin = -Inf, ymax = Inf, 
                fill =intervention_label),
            alpha=0.2) +
  geom_hline(aes(yintercept=0),lty=2) +
  geom_line(aes(x = date,percent_change_from_baseline, color=subregion)) +
  geom_smooth(aes(x = date,percent_change_from_baseline, color=subregion),
              span = 0.1) +
  facet_wrap(~field) +
  colorspace::scale_color_discrete_qualitative() +
  scale_x_date(date_breaks = "1 month",date_labels = "%b") +
  # theme(legend.position="bottom")
  labs(title = "Government interventions and Google mobility reports",
       subtitle = "Lima Metropolitan Area and Callao Region, Peru 2020",
       x = "Date",
       y = "% change from baseline",
       fill = "Interventions",
       color = "Region")
```



### Import initial stages of the WHO sitreps

- Created and maintained by [Fabienne Krauer](https://github.com/fkrauer/COVID-19)

```{r,fig.height=7.5,fig.width=9}
library(covid19viz)

# paste last update available at
# https://github.com/fkrauer/COVID-19
update <- "2020-03-10"

# apply
who_sitrep_country_report(
  update = update,
  country_region = "Brazil")
```

